\documentclass[a4paper, UTF-8]{article}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{tikz}
\usetikzlibrary{automata, positioning, arrows, trees}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{url}
\usepackage{booktabs}
\usepackage{nicefrac}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{amssymb}

\theoremstyle{plain}
\newtheorem{theorem}{\indent Theorem}[section]
\newtheorem{lemma}[theorem]{\indent Lemma}
\newtheorem{assumption}{\indent Assumption}[section]
\newtheorem{note}[theorem]{\indent Notation}
\newtheorem{proposition}[theorem]{\indent Proposition}
\newtheorem{corollary}[theorem]{\indent Corollary}
\newtheorem{definition}{\indent Definition}[section]
\newtheorem{example}{\indent Example}[section]
\newtheorem{remark}{\indent Remark}[section]
\newenvironment{solution}{\begin{proof}[\indent\bf Solution]}{\end{proof}}
\renewcommand{\proofname}{\indent\bf Proof}

\begin{document}

\title{\bf{Lab 7 Answers}}
\author{Yanqiao Chen*\\ 
  Department of Computer Science and Engineering\\
  Southern University of Science and Technology\\
  \texttt{12412115@mail.sustech.edu.cn}
  }
\maketitle

\section{Question 1} 

Informally, the PDA may first read $a^*b^*$, and push $a$ in the first stack, $b$ in the second stack. Then every time it reads $c$, it pops one $a$ and one $b$. Finally, if the input is empty and both stacks are empty, the PDA accepts the input. If one stack is empty but the other is not, i.e., it reads one $\mathdollar$ and an $a$ or $b$ from the top of the stacks, the PDA will not accept the input. If both stacks are not empty at the end of the input, the PDA will also not accept the input. We provide a formal one: 

\begin{proof} 
The 2-PDA recognizes $A = \{a^k b^k c^k \mid k \ge 0\}$ by using its two stacks to track the counts of different symbols in stages:
\begin{enumerate}
    \item \textbf{Stage 1 ($a^*$):} For every $a$ read from the input, push a symbol onto Stack 1.
    \item \textbf{Stage 2 ($b^*$):} For every $b$ read from the input, push a symbol onto Stack 2.
    \item \textbf{Stage 3 ($c^*$):} For every $c$ read from the input, pop one symbol from Stack 1 and one symbol from Stack 2 simultaneously.
\end{enumerate}
If the input is exhausted and both stacks are empty, the 2-PDA accepts; otherwise, it rejects. For $k=0$, the machine starts with empty stacks and accepts the empty string immediately.
\end{proof}

\section{Question 2}
The following trace follows the specific TM $M$ that decides $\{0^{2^n} \mid n \ge 0\}$.

\subsection{a) Input: 0} 
1. $q_1 0$ $\to$ 2. $\sqcup q_2 \sqcup$ $\to$ 3. $\sqcup \sqcup q_{accept}$

\subsection{b) Input: 00} 
1. $q_1 00 \to$ 2. $\sqcup q_2 0 \to$ 3. $\sqcup x q_3 \sqcup \to$ 4. $\sqcup q_5 x \to$ 5. $q_5 \sqcup x \to$ 6. $\sqcup q_2 x \to$ 7. $\sqcup x q_2 \sqcup \to$ 8. $\sqcup x \sqcup q_{accept}$

\subsection{c) Input: 000} 
1. $q_1 000 \to$ 2. $\sqcup q_2 00 \to$ 3. $\sqcup x q_3 0 \to$ 4. $\sqcup x 0 q_4 \sqcup \to$ 5. $\sqcup x 0 \sqcup q_{reject}$

\subsection{d) Input: 000000} 
1. $q_1 000000$ \\
2. $\sqcup q_2 00000$ \\
3. $\sqcup x q_3 0000$ \\
4. $\sqcup x 0 q_4 000$ \\
5. $\sqcup x 0 x q_3 00$ \\
6. $\sqcup x 0 x 0 q_4 0$ \\
7. $\sqcup x 0 x 0 x q_3 \sqcup$ \\
8. $\sqcup x 0 x 0 q_5 x$ (Start of leftward scan) \\
... (Return to start) ... \\
9. $\sqcup q_2 x 0 x 0 x$ \\
10. $\sqcup x q_2 0 x 0 x$ \\
11. $\sqcup x x q_3 x 0 x$ \\
12. $\sqcup x x x q_3 0 x$ \\
13. $\sqcup x x x 0 q_4 x$ \\
14. $\sqcup x x x 0 x q_4 \sqcup$ \\
15. $\sqcup x x x 0 x \sqcup q_{reject}$ (Rejects because it hits $\sqcup$ while in state $q_4$)

\section{Question 3} 
\begin{itemize}
    \item [a.] \textbf{Yes}. The blank symbol $\sqcup$ is a member of the tape alphabet $\Gamma$ by definition.
    \item [b.] \textbf{No}. The input alphabet $\Sigma$ cannot contain $\sqcup$, whereas $\Gamma$ must contain $\sqcup$. Thus, $\Sigma \subsetneq \Gamma$.
    \item [c.] \textbf{No}. In the standard formal definition, the transition function maps to $\{L, R\}$, meaning the head must move. (Stay-put $S$ is a variant).
    \item [d.] \textbf{No}. The definition requires that the accept state $q_{accept}$ and reject state $q_{reject}$ be distinct ($q_{accept} \neq q_{reject}$).
\end{itemize}

\section{Question 4}
To support $k=0$, we add a transition from $q_1$ to handle the empty string (reading $\sqcup$ initially).
\begin{center}
\begin{tikzpicture}[shorten >=1pt,node distance=3cm,on grid,auto, every state/.style={minimum size=1.2cm}]
    \node[state,initial]   (q1) {$q_1$};
    \node[state]           (q2) [right=of q1] {$q_2$};
    \node[state]           (q3) [right=of q2] {$q_3$};
    \node[state]           (q4) [below=of q3] {$q_4$};
    \node[state]           (q8) [above=of q2] {$q_8$};
    \node[state,accepting] (qa) [right=of q8] {$q_{\text{accept}}$};
    \path[->]
        (q1) edge node {$a\!\to\!X,R$} (q2)
             edge node[above left] {$Y\!\to\!Y,R$} (q8)
             edge[bend left=45] node {$\sqcup\!\to\!\sqcup,R$} (qa) 
        (q2) edge[loop above] node {$\begin{array}{c}a\!\to\!a,R\\Y\!\to\!Y,R\end{array}$} (q2)
             edge node {$b\!\to\!Y,R$} (q3)
        (q3) edge[loop above] node {$\begin{array}{c}b\!\to\!b,R\\Z\!\to\!Z,R\end{array}$} (q3)
             edge node[left] {$c\!\to\!Z,L$} (q4)
        (q4) edge[loop right] node {$\begin{array}{c}a\!\to\!a,L\\b\!\to\!b,L\\Y\!\to\!Y,L\\Z\!\to\!Z,L\end{array}$} (q4)
             edge[bend left=20] node[below left] {$X\!\to\!X,R$} (q1)
        (q8) edge[loop above] node {$\begin{array}{c}Y\!\to\!Y,R\\Z\!\to\!Z,R\end{array}$} (q8)
             edge node {$\sqcup\!\to\!\sqcup,R$} (qa);
\end{tikzpicture}
\end{center}

\section{Question 5}
\subsection{a)}

On input $w$ over $\{0,1\}$:
\begin{itemize}
    \item Scan the tape. If no unmarked $0$ or $1$ remains, \emph{accept}.
    \item Find the leftmost unmarked $0$ and mark it with $X$.
    \item Continue scanning right and find the leftmost unmarked $1$; mark it with $X$.
    \item If a matching $1$ cannot be found, \emph{reject}.
    \item Return the head to the leftmost unmarked symbol and repeat.
\end{itemize}
Each iteration cancels one $0$ with one $1$, so the machine accepts exactly when their counts are equal.

\subsection{b)}

On input $w$ over $\{0,1\}$:
\begin{itemize}
    \item Scan the tape. If no unmarked symbols remain, \emph{accept}.
    \item Find the leftmost unmarked $1$ and mark it with $X$.
    \item Scan right to find the first unmarked $0$ and mark it with $X$; then find a second unmarked $0$ and mark it with $X$.
    \item If two matching $0$'s cannot be found, \emph{reject}.
    \item Return the head to the leftmost unmarked symbol and repeat.
\end{itemize}
Each iteration consumes one $1$ and two $0$'s, so the machine accepts exactly when $\#0 = 2\cdot\#1$.

\subsection{c)}

This is the complement of the language in part~(b). Since Turing machines are closed under complement for \emph{decidable} languages, construct a TM $M'$ that runs the decider $M$ from part~(b) and then flips the answer:
\begin{itemize}
    \item Run $M$ on $w$.
    \item If $M$ accepts, $M'$ \emph{rejects}; if $M$ rejects, $M'$ \emph{accepts}.
\end{itemize}
Thus $M'$ decides $\{w \mid w \text{ does not contain twice as many 0s as 1s}\}$. 

\section{Question 6} 

\begin{proof}

Any Turing machine can simulate a Turing machine with left reset, thus Turing Machines $\Leftarrow$ Turing Machines with left reset. We then show the opposite direction, i.e., Turing Machines $\Rightarrow$ Turing Machines with left reset.

We use the extended alphabet $\Gamma' = \Gamma \times \{0, 1, 2\}$ and state with $Q' = Q \times \{normal, clear\}$. 

When we want to reset, we denote the symbol with $2$, i.e., $c^2$. Then we turn to the leftmost symbol and change it into $a^1$. Then we follow the cycles: 
\begin{itemize}
    \item Reset and read until we meet a symbol with $1$, read the next one, if it is denote with $2$, change it into $0$.
    \item If it is denote with $0$, change it into $1$ and reset with state changing into the clearing state. 
    \item We continue to read until we meet a symbol with $1$, change it into $0$ and reset with normal mode. 
    \item Repeat. 
\end{itemize}

After this, the left one is the symbol with $1$, and the right one is the symbol with $2$. Then we stop at the leftmost symbol with notation $1$ and change it into $0$. Then the operation is equivalent to the original Turing machine with left shift.

\begin{example}
On the tape 
$$
abba
$$
and we are at the end. We first denote the symbol with $2$, i.e.,
$$
abba^2
$$

For simplicity, we only explicitly write out symbol with $1$ and $2$.

Then we turn to the leftmost symbol and change it into $a^1$, i.e., $a^1bba^2$, then we follow the cycles:
\begin{enumerate}
    \item We find the symbol with $1$ and read the next one, which is $b^0$, we change it into $b^1$ and reset with state changing into the clearing state, i.e., $a^1b^1ba^2$.
    \item We find the symbol with $1$ and change it into $0$, i.e., $ab^1ba^2$, and reset with normal mode.
    \item We find the symbol with $1$ and read the next one, which is $b^0$, we change it into $b^1$ and reset with state changing into the clearing state, i.e., $ab^1b^1a^2$.
    \item We find the symbol with $1$ and change it into $0$, i.e., $abb^1a$, and reset with normal mode.
    \item We find the symbol with $1$ and read the next one, which is $a^2$, we change it and stop. Then we reset and read the symbol with $1$, i.e., $abb^1a$, when we read it, we do the operation after the left shift in the original Turing machine and our head is on the third $b$, which is equivalent to the original Turing machine with left shift.
\end{enumerate}

For any $\delta(q, a) = (q', b, L)$, we may use the operation above to simulate it, and the last step will be $\delta(q, a^1) = (q', b^0, D)$, where $D$ denotes an head operation. 

\end{example}

The Turing machine with left reset can simulate any Turing machine, thus Turing Machines $\Rightarrow$ Turing Machines with left reset.

This means that Turing Machines $\Leftrightarrow$ Turing Machines with left reset, thus it can recognize all the Turing-recognizable languages.

\end{proof}

\section{Question 7} 

Because the head can never move left, once it leaves a tape cell, the information stored there cannot be re-read. The machine's future decisions depend only on the current state and the remaining unread input. This behavior is equivalent to a Finite Automaton (DFA), where the tape acts as the input string.

Formally: 

\begin{theorem}
    A Turing machine whose transition function uses only \emph{Right} (\texttt{R}) and \emph{Stay} (\texttt{S}) operations recognizes exactly the class of regular languages.
\end{theorem}

\begin{proof}
    Let $M = (Q, \Sigma, \Gamma, \delta, q_0, q_{\text{accept}}, q_{\text{reject}})$ be a Turing machine whose transition function is of the form
    \[
        \delta \colon Q \times \Gamma \longrightarrow Q \times \Gamma \times \{R, S\}.
    \]
    We construct a DFA $D = (Q_D, \Sigma, \delta_D, q_{\text{start}}, F_D)$ such that $L(D) = L(M)$.

    \paragraph{Observation.}
    Since $M$ is never allowed to move its head left, any symbol written on a tape cell at position $i$ will never be read again after the head leaves position $i$. Consequently, the only effect of a sequence of $S$-transitions on the same cell is the state $M$ is in when it finally executes an $R$-transition.

    \paragraph{State set.}
    The states of $D$ are drawn from $Q$ together with a trap state:
    \[
        Q_D \;=\; Q \cup \{\,q_{\text{dead}}\,\}.
    \]

    \paragraph{Transition function $\delta_D$.}
    Fix a state $q \in Q$ and an input symbol $a \in \Sigma$.
    Starting from the configuration in which $M$ is in state $q$ and reading $a$ on the tape,
    repeatedly apply transitions of the form $\delta(\cdot, \cdot) = (\cdot, \cdot, S)$.
    There are two possibilities:
    \begin{enumerate}
        \item[(i)] The machine eventually applies a transition with $R$; let $p \in Q$ be the state entered immediately after that $R$.  Define
            \[
                \delta_D(q, a) \;:=\; p.
            \]
        \item[(ii)] The machine never applies an $R$-transition; it either loops forever through $S$-transitions or halts in $q_{\text{reject}}$.  Define
            \[
                \delta_D(q, a) \;:=\; q_{\text{dead}}.
            \]
    \end{enumerate}
    Finally, set $\delta_D(q_{\text{dead}}, a) := q_{\text{dead}}$ for all $a \in \Sigma$.

    \paragraph{Start state.}
    The initial state of $M$ need not itself be a state from which $M$ executes $R$.
    To handle the prefix of the computation before the first $R$-transition, define
    \[
        q_{\text{start}} \;:=\;
        \begin{cases}
            p & \text{if starting from }(q_0, \sqcup)\text{ the first }R\text{-transition leads to }p \in Q,\\
            q_{\text{dead}} & \text{otherwise.}
        \end{cases}
    \]
    (More symmetrically, one may augment the input with a left-endmarker and treat the initial move from $\sqcup$ uniformly.)

    \paragraph{Accepting states.}
    Once the input has been consumed, $M$ reads blank symbols $\sqcup$ to the right of the input.  It is in these post-input $S/R$ cycles that $M$ decides acceptance.  Hence define
    \[
        F_D \;=\; \{\, q \in Q \mid \text{starting from configuration }(q, \sqcup),\; M\text{ eventually enters }q_{\text{accept}} \,\}.
    \]

    \paragraph{Correctness.}
    We argue by induction on the tape position $i \geq 1$ that after reading the first $i$ input symbols, $D$ is in exactly the state $M$ occupies immediately after the $R$-transition that moves its head from cell $i$ to cell $i+1$.  The base case ($i=0$) holds by definition of $q_{\text{start}}$; the inductive step follows from the definition of $\delta_D$.  When the input is exhausted, $D$ is in the state $M$ would be in upon first reading $\sqcup$ after the last input symbol.  By construction, this state lies in $F_D$ iff $M$ accepts.

    Therefore $L(D) = L(M)$, and $M$ recognizes a regular language.
\end{proof}

Thus the machine is not equivalent to a Turing machine, and it cannot recognize all the Turing-recognizable languages. It can only recognize regular languages.


\section{Question 8} 
The collection of decidable languages is closed under:
\begin{itemize}
    \item \textbf{Union/Intersection:} For input $w$, run $M_1$ and $M_2$ sequentially on a multi-tape TM. For union, accept when one accepts, reject when all reject. For intersection, accept when all accept, reject when one rejects.
    \item \textbf{Complementation:} Run the decider $M_1$ and swap its $q_{accept}$ and $q_{reject}$ states. Or simulate this the machine $M_1$ on a TM $M_2$, if $M_1$ accepts, $M_2$ rejects; If $M_1$ rejects, $M_2$ accepts.
    \item \textbf{Concatenation/Star:} Construct a Non-deterministic Turing Machine (NTM) that non-deterministically partitions the input into substrings and verifies each substring using the respective deciders. Since any NTM can be simulated by a deterministic TM, the resulting language is decidable.
\end{itemize}

\section{Question 9}

\begin{proof}
Let $L_1$ and $L_2$ be two Turing-recognizable languages, and let $M_1$ and $M_2$ be TMs that recognize them, respectively.

\paragraph{a. Union.}
Construct a TM $M$ that on input $w$ runs $M_1$ and $M_2$ on $w$ in parallel (using dovetailing). If $M_1$ accepts, $M$ \emph{accepts}; if $M_2$ accepts, $M$ \emph{accepts}. If $w \in L_1 \cup L_2$, at least one of the two machines accepts, so $M$ accepts. Hence $L_1 \cup L_2$ is Turing-recognizable.

\paragraph{b. Concatenation.}
Construct a TM $M$ that on input $w$ nondeterministically guesses a split $w = uv$. It then runs $M_1$ on $u$ and $M_2$ on $v$. If both accept, $M$ \emph{accepts}. If $w \in L_1 L_2$, there exists at least one split on which both machines accept, so $M$ accepts. Thus $L_1 L_2$ is Turing-recognizable.

\paragraph{c. Star.}
Construct a TM $M$ that on input $w$ first \emph{accepts} if $w = \varepsilon$. Otherwise it nondeterministically guesses a decomposition $w = w_1 w_2 \cdots w_k$ with $k \ge 1$ and each $w_i \neq \varepsilon$, and runs $M_1$ on each piece. If every run accepts, $M$ \emph{accepts}. This shows $L_1^*$ is Turing-recognizable.
\end{proof}

\section{Question 10} 

\subsection{a)}

On input $x$ (a positive binary number with the least significant bit on the right):
\begin{itemize}
    \item Move to the rightmost bit.
    \item If it is $1$, change it to $0$ and halt.
    \item If it is $0$, change it to $1$ and move left; repeat moving left until finding the first $1$, change it to $0$, and halt.
\end{itemize}
This implements binary subtraction-by-one with borrowing.

\subsection{b)}

Assume the input is of the form $x \# y$. The TM computes $x - y$ as follows:
\begin{itemize}
    \item \emph{Loop:} Scan to the least significant bit of $y$.
    \item If all bits of $y$ are $0$ (i.e., $y$ has been reduced to zero), erase $y$ and the delimiter $\#$, compact the tape, and halt.
    \item Otherwise, apply the subroutine from part (a) to decrement $y$ by one.
    \item Then apply the same subroutine to decrement $x$ by one.
    \item Return to the start of the loop.
\end{itemize}
Since each iteration decrements both numbers by one, after exactly $y$ iterations $y$ becomes $0$ and $x$ has been reduced to $x - y$.

\subsection{c)}

Assume the input is $x \# y$ with $x > y$. The TM computes the quotient $q = \lfloor x / y \rfloor$ as follows:
\begin{itemize}
    \item Maintain a quotient area $q$ on the tape, initially $0$.
    \item Repeatedly subtract $y$ from the current $x$ using the procedure of part (b).
    \item Each time the subtraction succeeds, increment $q$ by one: scan to the least significant bit of $q$; flip trailing $1$'s to $0$ and the first $0$ to $1$; if all bits were $1$, append a new most significant $1$.
    \item If a subtraction fails because $x < y$, erase $x$, $y$, and the delimiter, leaving only $q$, and halt.
\end{itemize}
The loop executes exactly $\lfloor x / y \rfloor$ times, so the final $q$ equals the integer quotient.

\end{document}